\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-20 arch.tex
addressed kcomments the best I could

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on {\em external} state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable.% : \comment{Needed to say that
  % early on but the flow of this paragraph is not very tight}
\kcomment{Looks fine to me unless you want to clarify further.}

@@

Requests from a reactor can be {\em recalled} when it is determined
that they are not part of its projected plan anymore. For example if a
reactor needs to replan due to an unexpected event, it will recall all
requests it produced that have not been observed \kcomment{replace
  ``observed'' with ``not incorporated by the owner reactor''?}
yet. This recall is sent to the reactor that received the goal(s) to
remove it/them from its current set of objectives.

@@

Both these deliberative \kcomment{``deliberative??''; something else
  or remove altogether perhaps?} functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference for the current tick by
identifying the current value of {\em internal} state variables from
the latest updates on {\em external} state variables. This process
propagates up in the reactor dependency graph at every tick to ensure
that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process focused on the future evolution of state variables; it can
take multiple steps spanning multiple ticks for a reactor to produce a
plan that is both complete and valid. How the reactor identifies that
its plan is complete and valid is its responsibility and depends on
the way each reactor resolves its plan at each {\em step}. The general
idea is that the reactor is able to find a valid strategy in order to
satisfy as many of its {\em internal} goals as possible. For \eu, this
means that the solver is unable to identify any further flaws to be
resolved in the plan.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
stops, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps to
target the synchronization problem by ensuring that the new {\em
  external} state updates are compatible with this plan. If so,
synchronization propagates these events in the plan and can uses it as
an aid in identifying its {\em internal} timelines \kcomment{``use it
  as an aid to identify'' what though?} based on this plan.

If they are not compatible, then {\em synchronization} identifies a
plan inconsistency. This results in the retraction of decisions made
during the \kcomment{previous?} {\em deliberation} step. % in order to
% allow {\em synchronization} to identify its {\em internal} state.
Such a relaxation can be viewed as a backtracking % on the plan search
% triggered exogenously by the world evolution. It is the way our
allowing the reactor to identify that its current plan has failed to
execute triggering further deliberations steps to find an alternate
strategy.

@@

In \texttt{T-REX}'s \eu reactor this decision-making requirement is a
new type of flaw which forces identification \kcomment{identification
  of what? Probably need a little more explanation here.} of internal
state of the reactor for the current tick. This flaw can be defined
as:

@@

As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of unit decisions \kcomment{has a unit
  decision been explained in Europa?}. Since such a process is
required by the architecture at \emph{every} tick for \emph{every}
reactor the model needs to be designed carefully.

@@

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch
  window} from Definition \ref{def:dispatch} which filters out at the
agent level those goals that are before the reactor execution latency
or too far into the future. The planning window inside of a reactor
refines this filtering further, as tokens can be excluded dynamically
from deliberation as their insertion in the plan and scheduling with
other tokens generated, can reduce their temporal scope outside of the
planning scope \kcomment{ this last sentence should be reworked}. This
helps the deliberation solver reduce the number of tokens to be
evaluated reducing deliberation cost and increasing performance. As
time advances, flaws that were initially beyond this temporal horizon
will become active which a reactor can evaluate. This results in an
apparent \emph{continuous planning} of the reactor. When there are no
more flaws present in the current planning scope a partial plan
solution is found. This has two effects:

>>>>

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on {\em external} state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable.% : \comment{Needed to say that
  % early on but the flow of this paragraph is not very tight}
%\kcomment{Looks fine to me unless you want to clarify further.}

@@

Requests from a reactor can be {\em recalled} when it is determined
that they are not part of its projected plan anymore. For example if a
reactor needs to replan due to an unexpected event, it will recall all
requests it produced that have not been executed yet.
% \kcomment{replace
%   ``observed'' with ``not incorporated by the owner reactor''?}
% yet
 This recall is sent to the reactor that received the goal(s) to
remove it/them from its current set of objectives.

@@

Both these 
% deliberative \kcomment{``deliberative??''; something else
%   or remove altogether perhaps?} 
functions have different scope and temporal constraints. 
The \texttt{synchronize} call is considered as atomic and will 
focus only on state inference for the current tick by
identifying the current value of {\em internal} state variables from
the latest updates on {\em external} state variables. This process
propagates up in the reactor dependency graph at every tick to ensure
that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process focused on the future evolution of state variables; it can
take multiple steps spanning multiple ticks for a reactor to produce a
plan that is both complete and valid. How the reactor identifies that
its plan is complete and valid is its responsibility and depends on
the way each reactor resolves its plan at each {\em step}. The general
idea is that the reactor is able to find a valid strategy in order to
satisfy as many of its {\em internal} goals as possible. For \eu, this
means that the solver is unable to identify any further flaws to be
resolved in the plan.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
stops, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps to
target the synchronization problem by ensuring that the new {\em
  external} state updates are compatible with this plan. If so,
synchronization propagates these events in the plan and can uses it as
an aid in resolving its {\em current state flaw} by using the plan
information on the exepected {\em internal} state evolution nearby the
execution frontier. 
% \kcomment{``use it
%   as an aid to identify'' what though?}

If they are not compatible, then {\em synchronization} identifies a
plan inconsistency. This results in the retraction of decisions made
during previous {\em deliberation} steps. % in order to
% allow {\em synchronization} to identify its {\em internal} state.
Such a relaxation can be viewed as a backtracking % on the plan search
% triggered exogenously by the world evolution. It is the way our
allowing the reactor to identify that its current plan has failed to
execute triggering further deliberations steps to find an alternate
strategy.

@@

In \texttt{T-REX}'s \eu reactor this decision-making requirement is a
new type of flaw which requires to identify the value of all the {\em
  internal} state variables for the current tick.
% forces identification \kcomment{identification
%   of what? Probably need a little more explanation here.} of internal
% state of the reactor for the current tick. 
This flaw can be defined as:

@@

As the simple example above shows, synchronization involves a
deliberation process with the risk that the search for a solution is
not a straightforward sequence of staright-forward decisions where 
the resolution can be easily identified by the solver as any other
option would leafd immediately to an inonsistency in the plan.
% \kcomment{has a unit decision been explained in Europa?}. 
Since such a process is required by the architecture at \emph{every}
tick for \emph{every} reactor the model needs to be designed carefully.

@@

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch
  window} from Definition \ref{def:dispatch} which filters out at the
agent level those goals that are before the reactor execution latency
or too far into the future. The planning window reflect this filtering
of the agent by applying to tokens that are generated internally by the
reactor during its deliberation or the {\em goals} for which he
planner decided to postpone after the planning window -- probably
due to other tasks that needed completion beforehand.
% The planning window inside of a reactor
% refines this filtering further, as tokens can be excluded dynamically
% from deliberation as their insertion in the plan and scheduling with
% other tokens generated, can reduce their temporal scope outside of the
% planning scope \kcomment{ this last sentence should be reworked}. 
This allow the planner to focus on its planning horizon and report
decision on flawss thgat are beyond this scope on later ticks. As, 
time advance, these future flaws will eventuallty be part of the
planning window of the reactor and will be then resolved along with
potential new goals received from the agent.
% This helps the deliberation solver reduce the number of tokens to be
% evaluated reducing deliberation cost and increasing performance. As
% time advances, flaws that were initially beyond this temporal horizon
% will become active which a reactor can evaluate.
 This results in an apparent \emph{continuous planning} of the reactor. When there are no
more flaws present in the current planning scope a partial plan
solution is found. This has two effects:
